### 2017-04-20 LaTex/Paper.tex
Latex

%Preamble
\documentclass[a4paper,12pt]{article}
\usepackage[danish]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{todonotes}
\usepackage{textcomp}
\usepackage{caption}
\usepackage{pdfpages}
\usepackage{chngpage}
\usepackage{hyperref}
\usepackage{comment}
\usepackage{tikz}
\usepackage{braket}
\usepackage{hyperref}
\usepackage{float}
\usepackage{soul}
\usepackage{graphicx}
\usepackage{url}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{setspace}
\onehalfspacing

% Opening
\title{Organ donations}
\author{Emilie Nordby \& Simon Bjerg}
\date{20. April 2017}

\begin{document}
\maketitle
\newpage
\begin{abstract}

\end{abstract}	
\newpage
\tableofcontents

\newpage
\section{Indledning}

\subsection{Problemlegitimering}

\newpage
\begin{thebibliography}{9}

\bibitem{churn}
Daly, J. (2002), "Pricing for Profitability: Activity-Based Pricing for Competitive Advantage", John Wiley and Sons

\bibitem{infpre}
Breiman, L. (2001), "Statistical Modeling: The Two Cultures", Statistical Science, p 199-231

\bibitem{dong}
Om DONG Energy \\
"https://www.dongenergy.dk/om-dong-energy"

\bibitem{harv}
King, G. \& Zeng, L. (2001), "Logistic Regression in Rare Events Data"

\bibitem{conf}
Zar, J.H. (1984), "Biostatistical Analysis", Prentice Hall International New Jersey

\bibitem{old}
Forsberg, L., Wendelboe, L., Müller, C. \& Paldam, M. (2013), "Det digitale Danmark set med 65+ øjne", Teknologisk Institut

\bibitem{donglink}
Hvad består elprisen af? \\
"https://www.dongenergy.dk/privat/produkter-og-priser/el/spørgsmål-og-svar-om-el/hvad-består-elprisen-af"

\bibitem{moti}
McClelland, D. (1987), "Human Motivation", Cambridge University Press

\bibitem{smooth}
Barro, R. (2008), "Macroeconomics: A Modern Approach", Thomson South-Western

\bibitem{stat}
James, G., Witten, D., Hastie, T. \& Tibshirani, R. (2013), "An Introduction to Statistical Learning", Springer

\end{thebibliography}

\end{document}