### 2017-04-20 LaTex/Paper.tex
Ændret indledning

%Preamble
\documentclass[a4paper,12pt]{article}
\usepackage[danish]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{todonotes}
\usepackage{textcomp}
\usepackage{caption}
\usepackage{pdfpages}
\usepackage{chngpage}
\usepackage{hyperref}
\usepackage{comment}
\usepackage{tikz}
\usepackage{braket}
\usepackage{hyperref}
\usepackage{float}
\usepackage{soul}
\usepackage{graphicx}
\usepackage{url}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{setspace}
\onehalfspacing

% Opening
\title{Organ donations}
\author{Emilie Nordby \& Simon Bjerg}
\date{20. April 2017}

\begin{document}
\maketitle
\newpage
\begin{abstract}

Vores artikel handler om bla bla

\end{abstract}	
\newpage
\tableofcontents

\newpage
\section{Indledning}

Vores artikel handler om bla bla

\newpage
\begin{thebibliography}{9}

\bibitem{churn}
Daly, J. (2002), "Pricing for Profitability: Activity-Based Pricing for Competitive Advantage", John Wiley and Sons

\end{thebibliography}

\end{document}